\LARGE \bf Mean Estimation from Adaptive One-bit Measurements

### 2017-09-30 Writeups/proof of Lemma 7 for log-concave distributions.tex
addressed John's comments in V3, including an extention of Thm. 3 to MSE convergance

\begin{document}
\graphicspath{{../Figs/}}
\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
We consider the problem of estimating the mean of a normal distribution under the following constraint: the estimator can access only a single bit of each sample from this distribution. We study the squared error risk in this estimation as a function of the number of samples and one-bit measurements $n$. We consider an adaptive estimation setting where the single-bit sent at step $n$ is a function of both the new sample and the previous $n-1$ acquired bits. For this setting, we show that no estimator can attain asymptotic mean squared error smaller than $\pi/(2n)$ times the variance. 
%
In other words, one-bit restriction increases the number of samples required for a prescribed accuracy of estimation by a factor of at least $\pi/2$ compared to the unrestricted case.
{\color{red} This fundamental restriction on estimation in the adaptive setting implies a lower bound any distributed estimation setting from one-bit samples. }
%
In addition, we provide an explicit estimator that attains this asymptotic error, showing that, rather surprisingly, only $\pi/2<1.571$ times more samples are required in order to attain estimation performance equivalent to the unrestricted case. 
\end{abstract}

%{\color{red}  The $\mu$-sum problem (Kim-ElGamal Ch. 21) gives a lower bound. Can be reduced to the CEO.}

%{\color{red}  See LASSO results in http://arxiv.org/abs/1506.02181v1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:Intro}

Processing and estimating information from data collected at multiple physical locations is subject to communication constraints. 
%For example, emerging sensor network technologies in medicine and ``smart cities" use many low-cost sensors to collect large amounts of data and transmit it to remote locations. These sensors must operate under severe power restrictions, hence they are limited by their communication bandwidth. 
For example, consider large-scale sensor arrays where information is collected through multiple sensors and transmitted to a central estimation unit. In this scenario, the ability to estimate a particular parameter from data collected by the sensors is dictated not only by quality of observations and their numbers, but also by the available communication rate between the sensors and the central estimator. The question that we ask is to what extend a parametric estimation task is affected by communication constraints, and what are the fundamental performance limits in estimating a parameter subject to these restrictions. In this paper we answer this question in a particular simple setting: the estimation of the mean $\theta$ of a normal distribution with a known variance $\sigma^2$, under the constraint that only a single bit can be communicated on each sample $X_n$ from this distribution. As it turns out, the ability to share information among different samples before sending each single-bit message dramatically affects the performance in estimating $\theta$. We distinguish in particular among three different settings:
 \begin{itemize}
 \item[(i)] \emph{Centralized} encoding: all $n$ encoders confer and produce a single $n$ bit message which is a function of $X_1,\ldots,X_n$. 
 \item[(ii)] \emph{Adaptive} or \emph{sequential} encoding: the $n$th encoder observes $X_n$ and the $n-1$ previous single bit messages.
 \item[(iii)] \emph{Distributed} encoding: the output of the $n$th encoder is a single bit that is only a function of $X_n$.
 \end{itemize}
Clearly, as far as information sharing is concerned, settings (iii) is a more restrictive version of (ii) which is more restrictive than (i). We measure the estimation performance by the mean squared error (MSE) risk. Since without bit constraint this risk is equals $\sigma^2/n$, we are interested in particular in the \emph{asymptotic relative efficiency} (ARE) of estimators in the constrained setting, defined by the ratio between their MSE to $\sigma^2/n$ as $n$ goes to infinity. \\

In setting (i) the estimator can evaluate the sample mean and then communicate it using $n$ bits. This strategy leads to MSE behavior of $\sigma^2/n + o(2^{-n})$. Therefore, the ARE in this setting is $1$. Namely, asymptotically, there is no loss in performance due to the communication constraint under centralized encoding. In this work we show that a similar result does not hold even in setting (ii): the ARE of any adaptive estimation scheme is at least $\pi/2$. Namely, the single-bit per sample constraint incurs a minimal penalty of at least $1.57$ compared to an unconstrained estimator or to the optimal estimator in setting (i). In addition to this negative statement, we provide an estimator that almost surely attains this ARE. In other words, we show that the lower bound of $\pi/2$ on the asymptotic relative efficiency is tight, and that it is attained regardless of the particular realization of $\theta$ or the size of the parameter space from which it is taken. Clearly, the minimal penalty on the efficiency of $\pi/2$ also holds under setting (iii), although the question whether such efficiency is achievable (or otherwise, what is the minimal ARE) remains open.  \\

As the variance $\sigma^2$ goes to zero, the task of finding $\theta$ using one-bit queries in the adaptive setting (ii) is easily solved by a bisection style method over the parameter space. Therefore, the general case of non-zero variance is a reminiscent of the noisy binary search problem with possibly infinite number of unreliable tests \cite{cicalese2002least, Karp:2007:NBS:1283383.1283478}. However, since we assume a continuous parameter space, a more closely related problem is that of one-bit analog-to-digital conversion of a noisy signal. For example, sigma-delta modulator (SDM) analog-to-digital conversion \cite{1092194} uses one-bit threshold detector combined with a feedback loop to update an accumulated error state, and therefore falls under setting (ii). A SDM with a constant input $\theta$ corrupted by a Gaussian noise was studied in \cite{53738}, where it was shown that the output of the modulator converges to the true constant input almost surely. In other words, the SDM provides a consistent estimator for setting (ii). The rate of this convergence, however, was not analyzed and cannot be derived from the results of \cite{53738}. Our results imply that the rate of convergence of a SDM to a constant input is at most $\sigma^2\pi/2$ over the number of feedback iterations. \\

We note that even the centralized encoding of setting (i) poses a non-trivial challenge for the design and analysis of optimal encoding and estimation scheme. Indeed, the standard technique to encode an unknown random quantity using $n$ bits is attained by a scalar quantizer. However, the optimal design of this quantizer depends on the distribution of its input \cite{gray1998quantization}, which is the goal of our estimation task. Therefore a non-trivial exploration exploitation tradeoff arises. In our setting, the only missing parameter is the mean, which, under setting (i), is known to the encode with uncertainty interval proportional to $\sigma/\sqrt{n}$. Therefore, while it is clear that uncertainty due to quantization decreases exponentially in the number of bits, an exact expression for the MSE in this setting is not easy to derive. The situation is even more involved in the adaptive encoding of setting (ii): an encoding and estimation strategy that is globally optimal for $n-1$ adaptive one-bit messages of a sample of size $n-1$, may not lead to a globally optimal strategy upon the recipient of the $n$th sample. Conversely, any one-step optimal strategy, in the sense that it finds the best one-bit message as a function of the current sample and the previous $2^{n-1}$ messages, is not guaranteed to be globally optimal. Therefore, the main contribution of this paper is a lower bound on the MSE risk of any globally optimal strategy. This lower bound is obtained by showing that the Fisher information of any $n$ steps strategy is not larger than $2n/(\pi \sigma^2)$. The desired bound of $\pi \sigma^2/(2n)$ on the MSE follows from the van Trees version of the information inequality \cite{gill1995applications}. Finally, we show that an estimator that attains asymptotic MSE of $\sigma^2\pi/(2n)$ is obtained as a special case of \cite[Thm. 4]{polyak1992acceleration}. In addition to these two results, we also derive the one-step optimal strategy and demonstrates numerically that  the MSE under this strategy converge to $\pi \sigma^2/(2n)$. A proof for this converges remains elusive. \\

Our results imply that even under coarse quantization constraints, it is possible to achieve MSE in parametric estimation within only a relatively small penalty compared to the unconstrained estimator. A possible clue for this non-intuitive result is obtained from drawing the connection between our setting and the remote multiterminal source coding problem, also known as the CEO problem \cite{berger1996ceo, viswanathan1997quadratic, oohama1998rate, prabhakaran2004rate}. This connection, explained in details in Section~\ref{sec:ceo}, immediately leads to a lower bound of $4/3$ on the ARE in the distributed encoding of setting (iii). While this lower bound provides no new information compared to the lower bound of $\pi /2$ we derive here for setting (ii), it shows that the distributed nature of the problem is not a limiting factor in achieving MSE close to optimal even under one-bit quantization of each sample. \\

Setting (ii) and (iii) were considered in 
\cite{zhang2013information} under the more general assumption of $m$ machines each has access to $n/m$ independent samples. The main result of \cite{zhang2013information} are bounds on the estimation error as a function of the number of bits $R$ with which each machine can communicate. Specializing their result to our setting, by taking $m=n$ and $R=1$, leads to looser bounds then $\sigma^2\pi/(2n)$ for case (ii) and (iii). 
Other related works include statistical inference under multiterminal data compression \cite{han1987hypothesis, zhang1988estimation}, although their results apply only to setting (i). The counterpart of setting (iii) in the case of hypothesis testing was considered in \cite{52470}.
%
Estimation problem subject to the one-bit quantization constraint were also considered recently in compressed sensing \cite{boufounos20081,baraniuk2017exponential} and in MIMO detection in wireless communication \cite{singh2009limits}. \\

The rest of this paper is organized as follows: the main problem and notation are defined in Section~\ref{sec:problem}. In Section~\ref{sec:ceo} we illustrate a connection between our parametric estimation problem and the remote multiterminal lossy compression problem. Our main results are given in  Section~\ref{sec:sequential}, where we present a lower bound on estimation from one-bit samples in the adaptive scheme, as well as an achievable scheme that attains this bound. Concluding remarks are given in Section~\ref{sec:conclusions}. For sake of fluency, long proofs are deferred to the Appendix.

\section{Problem Formulation \label{sec:problem}}

\begin{figure}
\begin{center}
\begin{tikzpicture}[node distance=2cm,auto,>=latex]
  \node at (0,0) (source) {$X_1$} ;
  \node[int1, right of = source, node distance = 1.5cm] (enc1) {Enc 1};  
\draw[->,line width = 2pt] (source) -- (enc1);

 \node[below of = source, node distance = 1.5cm] (source2) {$X_2$};
\node[int1, right of = source2, node distance = 1.5cm] (enc2) {Enc 2};  
\draw[->,line width = 2pt] (source2) -- (enc2);

\node[below of = source2, node distance = 3cm] (source3) {$X_n$};
\node[int1, right of = source3, node distance = 1.5cm] (enc3) {Enc n};  
\draw[->,line width = 2pt] (source3) -- (enc3);

\node[above of = source, node distance = 1cm] (dist) {$X_i \sim {\mathcal N} \left(\theta, \sigma^2 \right)$};
\node[below of = source2, node distance = 0.5cm] {$\vdots$};

\node[int1, right of = enc2, node distance = 4cm ] (est) {Est};
\draw[->,line width = 2pt] (enc1) -| node[above, xshift = -1.5cm] (mes1) { $M_1 \in \left\{-1,1\right\}$} (est);   
\draw[->,line width = 2pt] (enc2) -- node[above, xshift = 0.5cm] (mes2) { $M_2 \in \left\{-1,1\right\}$} (est);   
\draw[->] (mes1) -- +(0,-0.7) -| (enc2);

\draw[->] (mes2) -- +(0,-0.7) -| +(-2.5,-1.2);
\draw[->] (mes2)+(-1.5,0.8) -- +(-1.5,-0.7);
\draw[->,line width = 2pt] (enc3) -| (est);

\node[below of = mes2, node distance = 2cm] (mes3) {$M_1,\ldots,M_{n-1} $};

\draw[->,line width = 2pt] (enc3) -| node[above, xshift = -2cm]  {$M_n \in \left\{-1,1\right\}$} (est);

\draw[->] (mes3) -- +(0,-0.5) -| (enc3);

\node[right of = est, node distance = 1.5cm] (dest) {$\widehat{\theta}_n$};
%         \node [int1] (dec) [right of=dest, node distance = 1.5cm,  align=center] {\small Dec };
%\node [int1] (enc) [right of = dec, node distance = 3cm]{Enc}; 
%\draw[->,line width=2pt] (dec) -- (dest);
\draw[->, line width=2pt] (est) -- (dest);
\end{tikzpicture}
\end{center}
\caption{\label{fig:sequential} Adaptive one-bit encoding: the $i$th encoder delivers a single bit message that is a function of its private sample $X_i$ and the previous $i-1$ messages $M_1,\ldots,M_{i-1}$.}
\end{figure}

Let $X_i$, $i=1,\ldots,n$, be $n$ independent samples from the normal distribution with mean $\theta$ and variance $\sigma^2$. We assume that the mean $\theta$ is drawn once from a prior distribution $\pi(\theta)$ on $\Theta$, which is a closed interval of the real line. We moreover assume that $\pi(\theta)$ is absolutely continuous with respect to the Lebesgue measure with density $\pi(d\theta)$. The problem we consider is the estimation of the parameter $\theta$ under the following constraints on the communication between the samples $X_1,\ldots,X_n$ and a centralized estimator: 
\begin{itemize}
\item[(i)] The estimator at time $n$ is only a function of the $n$ messages $M^n = \left(M_1,\ldots,M_n \right)$.
\item[(ii)] For each $i=1,\ldots,n$, the $i$th message $M_i$ is a function of the sample $X_i$ and the $i-1$ previous messages $M^{i-1}$.
\item[(iii)] The $i$th message $M_i$ takes only two possible values, say $1$ and $-1$. 
\end{itemize}
In other words, the $i$th message is defined by a function from the real line to $\{-1,1\}$ that is measurable with respect to the sigma algebra generated by $M^{i-1}$ and $X_i$, and the $n$ messages $M^n$ are the only available to the estimator. Upon observing $M^n$, the estimator produces an estimate $\widehat{\theta}_n(M^n)$ of $\theta$. A system describing the above scheme is illustrated in Fig.~\ref{fig:sequential}. \\

In this work we are concerned with the MSE risk defined as
\begin{equation}
\label{eq:error_def}
\mathbb E\left(\widehat{\theta}_n - \theta \right)^2,
\end{equation}
where the expectation is taken with respect to the distribution of $X^n$ and the prior distribution $\pi(\theta)$. \\
%It is well known that minmax estimation error 
%\begin{equation}
%E_n = \min \sup_{\theta \in \Theta}  \mathbb E\left[ \left(\widehat{\theta}_n - \theta \right)^2 | \theta \right],
%\end{equation}
%can be obtained from $\epsilon_n$ by assuming a particular least favorable prior distribution on the parameter space $\Theta$ \cite{casella1981}.

The main problem we consider is the minimization of \eqref{eq:error_def} over all encoding and estimation strategies, and the characterization of its minimal value as a function of  $n$. This minimization is the combination of the following two procedures: (1) selecting the $i$th message $M_i$ based on past messages and current observation $X_i$, and (2) estimating $\theta$ given messages $M^n$. We are interested in particular in the ARE of estimators of $\theta$ compared to the sample mean $\bar{X}_n = \frac{1}{n} \sum_{i=1}^n X_i$. Since $\sqrt{n}(\bar{X}_n-\theta)$ converges to $\mathcal N(0,\sigma^2)$, this ARE is defined as
\begin{equation}
{\sigma^2}  \mathbb E\left(\widehat{\theta}_n - \theta \right)^2. 
\label{eq:relative_efficiency}
\end{equation}
%{\color{red} Since the conditional expectation of $\theta$ with respect to $\bar{X}_n$ (and not $\bar{X}_n$) minimizes \eqref{eq:error_def}, we need to justify the fact that we use the sample mean as the reference estimator.}

In addition to the notations defined above, we denote by $\phi(x)$ the standard normal density and by $\Phi(x)$ the standard normal cumulative distribution function. \\%Derivatives with respect to $\theta$ are denoted by prime.\\

Before deriving our main results, we comment on the relation between our setting and the remote multiterminal source coding problem.

\section{Relation to Remote Multiterminal Source Coding \label{sec:ceo}}
In this section we draw a connection between our general problem of mean estimation from one-bit samples, to the remote multiterminal source coding problem \cite{berger1996ceo}, also known as the CEO problem. The setting of the CEO includes $n$ encoders, each has access to a noisy version of a random source sequence. 
% $\left\{ U_j,\,j=1,2,\ldots \right\}$. 
The $i$th encoder observes $k$ noisy source symbols and transmit $R_i k$ bits to a central estimator. \par
Assuming that $\theta$ is drawn randomly once from a prior $\pi(\theta)$, mean estimation from one-bit under distributed encoding (setting (iii) in the Introduction) corresponds to the CEO setting with block length $k=1$, where the $i$th encoder observes
\begin{equation}
\label{eq:Gaussian_channel}
X_i = \theta + \sigma Z_i,
\end{equation}
with $Z_1,\ldots,Z_n$ i.i.d. standard normal, and uses $R_i=1$ bits to transmit its observation. As a result, the quadratic distortion in the optimal source coding scheme for the CEO with $n$ terminals at rates $R_1 = \ldots = R_n = 1$ and Gaussian observation noise at each estimate of variance $\sigma^2$, provides a lower bound on the MSE distortion in estimating $\theta$ in the distributed encoding setting. 
 \\

A closed-form expression for the MSE distortion in the CEO is known only for the case where the source sequence is sampled independently from the normal distribution \cite{prabhakaran2004rate}. By using the characterization of this minimal distortion as the number of terminals goes to infinity, we conclude the following:
\begin{prop} \label{prop:ceo_lower_bound}
Assume that $\pi(\theta) = \mathcal N(0,\sigma_\theta^2)$. Then any estimator $\widehat{\theta}_n$ of $\theta$ in 
the distributed setting satisfies
\begin{equation} \label{eq:ceo_bound}
 n\mathbb E \left( \theta - \theta_n \right)^2 \geq \frac{4\sigma^2}{3} + o(1).
\end{equation}
\end{prop}

\begin{proof}
We consider the expression \cite[Eq. 10]{chen2004upper} that provides the minimal distortion $D^\star$ in the CEO with $L$ observers and under a total sum-rate $R_\Sigma = R_1 + \ldots +R_L$:
\begin{equation} \label{eq:ceo_optimal_sumrate}
R_{\Sigma} = \frac{1}{2} \log^+ \left[ \frac{\sigma_\theta^2}{D^\star} \left( \frac{D^\star L}{ D^\star L - \sigma^2 + D^\star \sigma^2 / \sigma_\theta^2 }\right)^L  \right].
\end{equation}
Assuming $R_\Sigma = n$ and $L=n$, we get
\begin{equation} \label{eq:ceo_optimal_sumrate}
n = \frac{1}{2} \log \left[ \frac{\sigma_\theta^2}{D^\star} \left(\frac{ D^\star n }{D^\star n - \sigma^2 + D^\star \sigma^2/\sigma_\theta^2 }  \right)^n  \right].
\end{equation}
The value of $D^\star$ that satisfies the equation above describes the MSE under an optimal allocation of the sum-rate $R_\Sigma = n$ among the $n$ encoders. Therefore, $D^\star$ provides a lower bound to the CEO distortion with $R_1=\ldots,R_n = 1$ and hence a lower bound to the minimal MSE in estimating $\theta$ in the distributed encoding setting. By considering $D^\star$ in \eqref{eq:ceo_optimal_sumrate} as $n\rightarrow \infty$, we conclude that 
\[
D^\star = \frac{ \frac{4\sigma^2}{3} }{n + \frac{4 \sigma^2}{3 \sigma_\theta^2} } + O(e^{-n}) =  \frac{4\sigma^2}{3n} + o(n^{-1}). 
\]
\end{proof}
Since the lower bound \eqref{eq:ceo_bound}  was derived assuming optimal allocation of the $n$ bits among the encoders, it may seem as if a tighter bound can be obtained in our case by considering the CEO distortion with $R_1=\ldots,R_n = 1$. However, an upper bound for the CEO distortion under this condition follows from \cite[Prop. 5.2]{KipnisRini2017}, and leads to
\[
D_{CEO} \leq  \left( \frac{1}{\sigma_\theta^2} +  \frac{3n}{4\sigma^2 + \sigma_\theta^2} \right)^{-1}   =
\frac{4 \sigma^2}{3n} +  \frac{\sigma_\theta^2}{3n} + o(n^{-1}),
\]
which is equivalent to \eqref{eq:ceo_bound} when $\sigma_\theta$ is small. \par
We conclude therefore that the difference between the MSE lower bound \eqref{eq:ceo_bound} and the actual MSE in the distributed encoding setting, is attributed exclusively to the ability to perform coding over blocks. Namely, the ability to consider $k>1$ independent realizations of $\theta$ versus only one in ours. In other words, it is the ability to exploit the geometry of a high-dimensional product space, rather than the distributed nature of the problem, that distinguishes between the CEO distortion and the mean estimation from one-bit samples.\par
In the next section we show that the ARE in adaptive encoding setting does not exceeds $\pi/2$, and thus provides a tighter lower bound for the distributed encoding setting than $4/3$ of  \eqref{eq:ceo_bound}.

\section{Results \label{sec:sequential}}
The first main result of this paper, as described in Thm.~\ref{thm:adpative_lower_bound} below, states that the ARE of any adaptive estimator cannot be lower than $\pi/2$. Next, we provide a particular adaptive estimation scheme and show in Thm.~\ref{thm:sgd} that its efficiency is $\pi/2$. Finally, in Thm.~\ref{thm:opt_one_step}, we provide an adaptive estimation scheme that is one-step optimal in the sense that at each step $i$, the message $M_i$ that minimizes the MSE given $X_i$ and the previous $M^{i-1}$ messages is chosen. While it is not clear whether the efficiency of this last scheme is $\pi/2$, numerical simulations suggests that the MSE of this scheme times $n$ also converges to $\pi/2$.

%\begin{figure}
%\begin{center}
%\begin{tikzpicture}[node distance=2cm,auto,>=latex]
%  \node at (0,0) (theta) {$\theta$} ;
%  \node at (2,0) (x1) {$X_1$};
%  \node[below of = x1, node distance = 1cm] (x2) {$X_2$};
%    \node[below of = x2, node distance = 1cm] (dots) {$\vdots$};
%  \node[below of = dots, node distance = 1cm] (xn) {$X_n$};
%  
%   \node[right of = x1, node distance = 2cm] (m1) {$M_1$};
%  \node[right of = x2, node distance = 2cm] (m2) {$M_2$};
%    \node[right of = xn, node distance = 2cm] (mn) {$M_n$};
%
%\draw[->] (theta) -- (x1);
%\draw[->] (theta) -- (x2);
%\draw[->] (theta) -- (xn);
%
%\draw[->] (x1)--(m1);
%\draw[->] (x2)--(m2);
%\draw[->] (xn)--(mn);
%
%\draw[->] (m1)--(m2);
%\draw[->, dashed] (m2)--(mn); 
%  
%\end{tikzpicture}
%\end{center}
%\caption{\label{fig:sequencial_graphical} Graphical model notation of the sequential setting.}
%\end{figure}

%\subsection{Properties of an Optimal Scheme}
%The sequential nature of the encoding and the estimation is best explain by the notions of actions and rewards taken from dynamic programming. The \emph{action} taken by the $i$th encoder is the design of a detection region $M_i^{-1}(1)$ that would lead to $M_i = 1$ assuming when $X_i \in M_i^{-1}(1)$ or $M_i=-1$ otherwise. The \emph{reward} (or penalty) is only evaluated at the end of the time horizon $n$.
%% Dynamic program. State = $M^n$. 
%% Reward is collected at the end
%%

\subsection{A lower bound on adaptive one-bit schemes}
Our first results asserts that the ARE \eqref{eq:relative_efficiency} of any adaptive estimation scheme is bounded from below by $\pi/2$, as follows from the following theorem:
\begin{thm}[minimal relative effeciency] \label{thm:adpative_lower_bound}
Let $\widehat{\theta}_n$ be any estimator of $\theta$ in the adaptive setting of Fig.~\ref{fig:sequential}.
% Let $\pi(\theta)$ be an absolutely continuous prior distribution on $\Theta$ with density function converges to zero at the endpoints of the interval $\Theta$. 
Assumes that $\pi(\theta)$ converges to zero at the endpoints of the interval $\Theta$. Then
\[
\mathbb E\left[ (\theta-\theta_n)^2 \right] \geq  \frac{\pi \sigma^2 }{2n +\pi \sigma^2  I_0}  =  \frac{\pi}{2n}\sigma^2+o(n^{-1}),
\]
where 
\[
I_0 = \mathbb E \left( \frac{d}{d\theta} \log \pi (\theta) \right)^2
\]
is the Fisher information with respect to a location model in $\theta$. 
\end{thm}

\subsubsection*{Sketch of Proof}
The main idea in the proof is to bound from above the Fisher information of any set of $n$ single-bit messages with respect to $\theta$. Once this bound is achieved, the result follows by using the van-Trees inequality \cite[Thm. 2.13]{tsybakov2008introduction},\cite{gill1995applications} which bounds from below the MSE of any estimator of $\theta$ by the inverse of the expected value of the aforementioned Fisher information plus $I_0$. The details are given in the Appendix.\\

>>>>

\begin{document}

\begin{lem} \label{lem:bound_intervals}
Assume that $X_1,\ldots,X_n$ are distributed with CDF $F(x)$ and PDF $f(x)$ such that $F(-x) = 1-F(x)$ (and $f(-x)=f(x)$). Assume moreover that 
$ f(x) = e^{g(x}}$ where $g(x)$ is concave and twice differentiable such that $g''(x)<0$. Then for any $x_1 \geq \ldots \geq x_n \in \mathbb R$, we have 
\begin{equation}
\frac{ \left(  \sum_{k=1}^n (-1)^{k+1} f(x_k) \right)^2} 
{\left( \sum_{k=1}^n (-1)^{k+1} F(x_k) \right)\left(1- \sum_{k=1}^n (-1)^{k+1} F(x_k) \right)  } \leq \frac{2} {\pi}. \label{eq:bound_intervals}
\end{equation}
\end{lem}

\subsubsection*{Proof of Lem.~\ref{lem:bound_intervals}}We use induction on $n \in \mathbb N$. For the base case $n=1$ we have 
\begin{equation} \label{eq:induction_base}
\frac{  f^2(x)} 
{F(x) \left(1 - F(x) \right) }.
\end{equation}
Since $f(x)$ is log-concave, so does $F(x)$. We now write $F(x) = e^{h(x)}$ with $h(x)$ concave. Note that $f(x) = h'(x) e^{h(x)}$, $h'(x)\geq 0$, $f(x) = f(-x) = h'(-x)e^{h(-x)} $, and $h''(x)\leq 0$ ($h(x)$ is twice differentiable since $F(x)$ is twice differentiable). 
Taking the logarithm of \eqref{eq:induction_base} leads to 
\begin{align*}
& 2 \log f(x) - \log F(x) + \log F(-x) \\
& = 2 \log h'(x) + h(x) + h(-x).
\end{align*}
Since $h'(x)$ is non-increasing, the maximal value of \eqref{eq:induction_base} is obtained as 
{HERE!! \color{red} finish proof by showing that \eqref{eq:induction_base} is smaller than $4f(0)$.\\
}
\\
Assume now that \eqref{eq:bound_intervals} holds for all integers up to some $n = N-1$ and consider the case $n = N$. The maximal value of \eqref{eq:bound_intervals} is attained for the same $(x_1,\ldots,x_N) \in \mathbb R^N$ that attains the maximal value of 
\begin{align*}
& g(x_1,\ldots, x_N) \triangleq 2 \log \left(  \sum_{k=1}^{N} (-1)^{k+1} f(x_k) \right) - \\
&  \log
\left( \sum_{k=1}^N (-1)^{k+1} F(x_k) \right)
-\log \left(1 -  \sum_{k=1}^N (-1)^{k+1} F(x_k) \right) \\
& = 2 \log \delta_N - \log \Delta_N - \log \left(1 - \Delta_N  \right),
\end{align*}
where we denoted $\delta_N \triangleq \sum_{k=1}^{N} (-1)^{k+1} f(x_k)$ and $\Delta_N =  \sum_{k=1}^N (-1)^{k+1} F(x_k)$. The derivative of $g(x_1,\ldots,x_N)$ with respect to $x_k$ is given by
\[
\frac{\partial  g}{\partial x_k} = \frac{2 (-1)^{k+1} f'(x_k)}{\delta_N} -\frac{(-1)^{k+1} f(x_k)}{\Delta_N } + \frac{(-1)^{k+1} \phi(x_k)}{1-\Delta_N }.
\]
We conclude that the gradient of $g$ vanishes only if
\[
\frac{f'(x_k)}{f(x_k)} = g'(x_k) = \frac{\delta_N}{2} \left( \frac{1}{\Delta_N} - \frac{1}{1-\Delta_N} \right),\quad k=1,\ldots,N.
\]
Due to our assumption that $g''(x) < 0$, the function $g'(x)$ is strictly decreasing. As a result, $g'(x)$ is one to one and the condition above implies $x_1 = \ldots = x_N$. If $N$ is odd then for $x_1=\ldots =x_N$ we have that the LHS of \eqref{eq:bound_intervals} equals
\[
\frac{f(x_1)^2}{ F(x_1) (1 - F(x_1))},
\]
which was shown to be smaller than $4f^2(0)$. If $N$ is even, then for any constant $c$ the limit of the LHS of \eqref{eq:bound_intervals} as $(x_1,\ldots,x_N)\rightarrow (c,\ldots,c)$ exists and equals zero. Therefore, the maximum of the LHS of \eqref{eq:bound_intervals} is not attained at the line $x_1=\ldots=x_N)$. We now consider the possibility that the LHS of \eqref{eq:bound_intervals} is maximized at the borders, as one or more of the coordinates of $(x_1,\ldots,x_N)$ approaches plus or minus infinity. For simplicity we only consider the cases where $x_N$ goes to minus infinity or $x_1$ goes to plus infinity (the general case where the first $m$ coordinates goes to infinity or the last $m$ to minus infinity is obtained using similar arguments). Assume first $x_N \rightarrow -\infty$. Then the LHS of \eqref{eq:bound_intervals} equals
\begin{align*}
\frac{ \left(  \sum_{k=1}^{N-1} (-1)^{k+1}f(x_k) \right)^2} 
{\left( \sum_{k=1}^{N-1} (-1)^{k+1} F(x_k) \right)\left(1- \sum_{k=1}^{N-1} (-1)^{k+1} F(x_k)  \right) } ,
\end{align*}
which is smaller than $2/\pi$ by the induction hypothesis. Assume now that $x_1 \rightarrow \infty$. Then the LHS of
\eqref{eq:bound_intervals} equals
\begin{align*}
& \frac{ \left(  \sum_{k=2}^{N} (-1)^{k+1}f(x_k) \right)^2} 
{\left( 1 + \sum_{k=2}^{N} (-1)^{k+1} F(x_k) \right)\left(1- 1 - \sum_{k=2}^{N} (-1)^{k+1} F(x_k)  \right) }  \\
& = \frac{ \left(  -\sum_{m=1}^{N} (-1)^{m+1}f(x'_m) \right)^2} 
{\left( 1 - \sum_{m=1}^{N-1} (-1)^{m+1} F(x'_{m}) \right)\left( \sum_{m=1}^{N-1} (-1)^{m+1} F(x'_{m})  \right) },
\end{align*}
where $x'_{m} = x_{m+1}$. The last expression is also smaller than $4f(0)$ by the induction hypothesis. This proves Lem.~\ref{lem:bound_intervals}. \\

\subsubsection*{Proof of Lem.~\ref{lem:fisher_bound}}
The Fisher information of $M$ with respect to $\theta$ is given by
\begin{align}
I_\theta & =  \mathbb E \left[ \left( \frac{d}{d\theta} \log P\left( M | \theta \right) \right)^2 |\theta \right] \nonumber \\
& = \frac{ \left(\frac{d}{d\theta} P(M=1|\theta) \right)^2}{P(M=1| \theta)} + \frac{ \left(\frac{d}{d\theta} P(M=-1|\theta) \right)^2} {P(M=-1| \theta)} \nonumber \\
& =  \frac{ \left( \frac{d}{d\theta} \int_A \phi \left( \frac{x-\theta}{\sigma} \right)dx \right)^2} { P(M=1| \theta) } + \frac{ \left( \frac{d}{d\theta}\int_A \phi \left( \frac{x-\theta}{\sigma} \right)dx \right)^2} { P(M=-1| \theta) } \nonumber \\ 
& \overset{(a)}{=} \frac{ \left( - \int_A \phi' \left( \frac{x-\theta}{\sigma} \right)dx \right)^2} {\sigma^2 P(M=1| \theta) } + \frac{ \left(- \int_A \phi' \left( \frac{x-\theta}{\sigma} \right)dx \right)^2} { \sigma^2P(M=-1| \theta) } \nonumber \\ 
& = \frac{\left( \int_A \phi'\left( \frac{x-\theta}{\sigma} \right) dx \right)^2 }{  \sigma^2 P(M=1 | \theta) \left(1-P(M=1|\theta) \right)  }, \nonumber \\
& = \frac{\left( \int_A \phi'\left( \frac{x-\theta}{\sigma} \right) dx \right) \left( \int_A \phi'\left( \frac{x-\theta}{\sigma} \right) dx \right)}{  \sigma^2 \left( \int_A \phi \left( \frac{x-\theta}{\sigma} \right) dx \right)  \left(1- \int_A \phi \left( \frac{x-\theta}{\sigma} \right) dx \right) }, \label{eq:lem_fisher_bound_proof1}
\end{align}
where differentiation under the integral sign in $(a)$ is possible since $\phi(x)$ is differentiable with absolutely integrable derivative $\phi'(x) = -x\phi(x)$. Regularity of the Lebesgue measure implies that for any $\epsilon>0$, there exists a finite number $k$ of disjoint open intervals $I_1,\ldots I_k$ such that 
\[
\int_{A\setminus \cup_{j=1}^k I_j }  dx < \epsilon \sigma^2,
\]
which implies that for any $\epsilon' > 0$, the set $A$ in \eqref{eq:lem_fisher_bound_proof1} can be replaced by a finite union of disjoint intervals without increasing $I_\theta$ by more than $\epsilon'$. It is therefore enough to proceed in the proof assuming that $A$ is of the form
\[
A = \cup_{j=1}^k (a_j,b_j),
\]
with $\infty \leq a_1 \leq \ldots a_k$, $b_1 \leq b_k \leq \infty$ and $a_j \leq b_j$ for $j=1,\ldots,k$. Under this assumption we have
\begin{align*}
\mathbb P(M_n=1| \theta) & = \sum_{j=1}^k \mathbb P\left(X_n \in (a_j,b_j) \right)  \\
& = \sum_{j=1}^k \left( \Phi \left(\frac{b_j-\theta}{\sigma} \right) -  \Phi \left(\frac{a_j-\theta}{\sigma} \right)  \right),
\end{align*}
so \eqref{eq:lem_fisher_bound_proof1} can be rewritten as
\begin{align}
& =   \frac { \left( \sum_{j=1}^{k} \phi \left(\frac{a_j-\theta}{\sigma} \right) - \phi \left( \frac{b_j-\theta} {\sigma} \right)  \right)^2 } 
{\sigma^2 \left( \sum_{j=1}^k \Phi \left( \frac{b_j-\theta }{\sigma}\right) - \Phi \left( \frac{a_j-\theta }{\sigma}\right)  \right) }  \nonumber \\
& \times \frac {1} 
{1- \left( \sum_{j=1}^k \Phi \left( \frac{b_j-\theta }{\sigma}\right) - \Phi \left( \frac{a_j-\theta }{\sigma}\right)  \right) } 
\label{eq:lemma_J}
\end{align}
It follows from Lem.~\ref{lem:bound_intervals} that for any $\theta \in \mathbb R$ and any choice of the intervals endpoints, \eqref{eq:lemma_J} is smaller than $2/(\sigma^2 \pi)$. Therefore, the proof of  Lem.~\ref{lem:fisher_bound} is now completed.
